---
title: "Proposal title"
subtitle: "Proposal"
format: html
editor: visual
---

```{r}
#| label: load-pkgs
#| message: false
library(tidyverse)
```

## Dataset

```{r}
#| label: load-dataset
#| message: false
```

A brief description of your dataset including its provenance, dimensions, etc. as well as the reason why you chose this dataset.

Make sure to load the data and use inline code for some of this information.

## Questions

The two questions you want to answer.

## Analysis plan

-   A plan for answering each of the questions including the variables involved, variables to be created (if any), external data to be merged in (if any).

## Plan of Attack

|             Week             |                    Weekly Tasks                    | Team members involved |
|:---------------------:|:-------------------------:|:---------------------:|
|     Till November 8^th^      |   Explore the dataset and the problem statement    |       Everyone        |
|              \-              |               Complete the proposal                |       Everyone        |
|     Nov. 9^th^ - 14^th^      |             Exploratory Data Analysis              |          TBD          |
|              \-              | Data cleaning and Data pre-processing based on EDA |          TBD          |
|     Nov. 15^th^ - 21^st^     |                 Data Visualization                 |          TBD          |
|              \-              |                Training and Testing                |          TBD          |
|     Nov. 22^nd^ - 29^th^     |      K-Nearest Neighbors Model implementation      |          TBD          |
|              \-              |         Random Forest Model implementation         |          TBD          |
| Nov. 30^th^ - December 7^th^ |                  Model Evaluation                  |          TBD          |
|              \-              |                  Model Comparison                  |          TBD          |
|     Dec. 8^th^ - 11^th^      | Refining the code with comments and final changes  |          TBD          |
|              \-              |     Write-up and presentation for the project      |          TBD          |
